Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-05-17 submission5/core.tex
Priority twos done

Two (compatible) states are variants with respect to a set of elements \(\components\)
if they agree on the value of every element not in \(\components\).
Our notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements.

@@

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if \(\mach\sim\nach\) and
\(\Delta(\mach,\nach) \subseteq \components\). \sna{define \(\sim\). Priority 2.}
%if all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
%and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).

@@

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose Coq for our implementation
to ensure that our definitions are well-formed \rb{, to reason about them,} and to leverage the
power of the QuickChick
testing framework~\cite{Pierce:SF4}.

\sna{Discuss false positives, false negatives, and the degree to which
  testing can increase our confidence. Priority 2.}

\subsection{Test Generation}
\sna{Restore Leo's old test gen discussion, then ping Leo if it seems that more
  is needed. Priority 2.}

To use QuickChick, we developed random test-case generators that produce
(1) an initial RISC-V machine state, including the program to be executed;
(2)
  an initial policy state, tagging instructions and the stack regions appropriately
  for the micro-policy (see \cref{sec:enforcement}); and
(3)
  a map from instructions to lists of their operations in the security semantics.
%
\rb{((say a few more words about 3?))}
%
\rb{
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also the policy
meta-information and code annotations that would normally be produced
by the compiler.
}

To write such generators we build on the work of Hri\c{t}cu et
al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution}, a technique that produces
programs that lead to longer executions---and hopefully towards more
interesting behaviors as a result.
%
\rb{
Generation by
execution receives as an input a partially instantiated machine state
and attempts to generate an instruction (or a sequence of instructions
such as a blessed sequence) that makes sense locally (e.g., jumps go
to a potentially valid code location and loads read from a
potentially valid stack location). Then we step the machine and repeat
the process until we generate or execute some target number of
instructions, or reach a point where the machine cannot step
any more.
}

We extend their technique with an additional
degree of statefulness:
\rb{
we keep track of the control flow behavior of
the program being generated. Each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points.
}
When complex calling conventions need to be
observed to avoid early failstops, extra care must be taken to ensure
that % any such
calls are well behaved. For example, immediately
after a call we need to initialize any stack-allocated variables
with an increased probability. However, at the same time, we need to
\rb{allow the generation to sometimes relax those constraints,
introducing potentially ill-formed flows:}
% allow for the possibility of,
for example, {\em not} performing such an initialization
\rb{or \emph{not} returning to the proper call site, so as}
to allow for potential attack vectors to manifest.
\rb{
This causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.}

>>>>

Two (compatible) states are variants with respect to a set of elements \(\components\)
if they agree on the value of every element not in \(\components\).
Our notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements.
Recall that \(\sim\) is a policy-specific compatibility relation.

@@

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if \(\mach\sim\nach\) and
\(\Delta(\mach,\nach) \subseteq \components\).
%if all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
%and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).

@@

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose Coq for our implementation
to ensure that our definitions are well-formed \rb{, to reason about them,} and to leverage the
power of the QuickChick
testing framework~\cite{Pierce:SF4}.
%
QuickChick is a randomized property-based testing framework. We use it to generate
random programs, then execute them and test that they fulfill our criteria.

Such testing is sound---it will not produce false positives---but necessarily
necessarily incomplete. We might test a flawed policy but fail to find a
program that exploits the flaw. Additionally, detecting violations of noninterference-style
properties is dependent on choosing appropriate variant states, so it is possible
to find a dangerous program but have it pass the test due to variant selection.
We can increased our confidence in our test coverage by {\em mutation testing},
in which we intentionally inject flaws into our policies and demonstrate that our testing
can find them.

\subsection{Test Generation}

To use QuickChick, we developed random test-case generators that produce
an initial RISC-V machine state tagged appropriately for the micro-policy
(see \cref{sec:enforcement}), including a code region containing a
low-level program. They must also produce the meta-information about
how instructions in that program map to security-relevant operations,
information that would normally be provided by the compiler.

To write such generators we build on the work of Hri\c{t}cu et
al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution}, a technique that produces
programs that lead to longer executions---and hopefully towards more
interesting behaviors as a result.

Each step of generation by execution takes a partially instantiated
machine state and attempts to generate an instruction
that makes sense locally (e.g., jumps go
to a potentially valid code location and loads read from a
potentially valid stack location). The generator repeats this process
for an arbitrary number of steps, or until it reaches a point where
the machine cannot step any more. Each time it generates a call or return,
it places the appropriate policy tags on the relevant instruction(s)
and records the operation.

We extend Hrit\c{t}cu et al.'s technique with additional statefulness
to avoid early failstops. For example, immediately
after a call we initialize any stack-allocated variables
with an increased probability. However, periodically, the generator
will relax those constraints and generate potentially ill-formed
code, such {\em not} performing such an initialization, writing outside
of the current stack frame, or attempting an ill-formed return sequence,
to allow for potential attack vectors to manifest.
